---
title: "Curating Graebner OSU dataset"
output:
  minidown::mini_document:
    code_folding:
      source: hide
      output: hide
      message: show
      warning: show
      error: show
    results_folding: hide #show, hide or none
    framework: water
    theme: light
    toc: true
    toc_float: true
    toc_highlight: true
---
```{r setup, eval = FALSE, include = FALSE}
setwd(here::here())
rmarkdown::render('./curation-rmd/Graebner/scripts/1_graebner_curation.Rmd', 
                  output_dir = './curation-rmd/Graebner/output/graebner_curation')
```
Information/ links for more information about the program and data:
https://agsci.oregonstate.edu/kbrec/crops/small-grains
https://agsci.oregonstate.edu/mes/crops/wheat
https://agsci.oregonstate.edu/system/files/x2019_ontario_elite_wheat_trial_2018-19_trial_26june2020.pdf
https://docs.google.com/spreadsheets/d/16C8GGw-vlITDhpl_ldUWZeJIFMvSwe1cTHmqI2aEQao/edit#gid=571577693

## Specify dependencies and file structure:
Note!! knitroutputfolder name should match rmarkdown::render output_dir >>>>

### Graebner files
```{r}
library(here)
source(here("R", "dependencies.R"))
main_folder <- "curation-rmd/Graebner"
knitroutputfolder <- here(main_folder, "output/graebner_curation")
data_folder <- here(main_folder, "data-raw")
auxiliary_files <- here(main_folder, "auxiliary_files")
#cultivar_aux_files <- here(auxiliary_files, "neely_cultivar")
#cultivar_data_folder <- here("controlled_vocab", "cultivar_original", "Neely")
```
### Database files
Note: readin.db() won't run properly if files are open in excel
```{r}
db <- readin.db()
db_metadata<- list.db_var("trials_metadata")
```

## Manual edits:

## Read in raw data:
### Generate a list of all of the files and sheets
If header is not in the first row, 
write.csv(sheet_info...) and manually fill in the 
number of rows to skip in each sheet.  See read.multsheets() for more
information.
```{r}
files <- list.files(data_folder, recursive = TRUE, pattern = ".csv")

sheet_info <- data.frame(filename = files) %>% 
  filter(!str_detect(filename, "curation_files")) %>%
  mutate(list_names = basename(filename)) %>%
  mutate(list_names = str_remove(list_names, ".csv")) %>%
  mutate(year = parse_number(filename)) %>%
  mutate(list_names = ifelse(!str_detect(list_names, '[0-9]'), paste(list_names, year, sep = "_"), list_names))
```
### Use read.multsheets()
```{r}
data_raw1 <- read.multsheets(data_folder, sheet_info, na = c("NA"), col_names = TRUE)
data_raw2 <- imap(data_raw1, function(x, y){ x %>% mutate(sourcefile = y)})

save(data_raw2,
     file = here(knitroutputfolder, "data_raw2.RData"))
```
### Remove duplicate columns
Use rm.dup_col() because it can detect and remove duplicates when:
- column names are not exact, column contents are
- if the only difference between two column contents are that one has some NAs,
then that column is removed and the other is retained
```{r}
data_raw3 <- imap(data_raw2, function(x,y){
    print(y)
    rm.dup_col(x, FALSE) %>% 
    select(where(~!all(is.na(.))))
})

```

## Curate column names:
### Get column names
```{r}
colnames <- summarize.variables(data_raw3, reg_ex = NULL)
colnames_summary <- colnames %>%
  group_by(colname) %>%
  summarize(across(everything(), ~ paste(na.omit(unique(.x)), collapse = "; ")))

write.csv(colnames_summary, here(knitroutputfolder, "colnames_summary.csv"), row.names = FALSE)
```
### Rename columns using auxiliary files
First, understand what each column means (description and units) by:

1. checking through annual reports

2. checking with Julia

3. lastly, asking the PI

Then, fill out the auxiliary file with the appropriate controlled vocabulary.

If there isn't a corresponding controlled vocabulary, then need to:

1. decide if the column can be captured by current controlled vocabulary. If so,
standardize to the controlled vocabulary

2. decide whether we need to keep the variable at all (can fill in auxiliary file
with "rm_col" in the controlled_vocab column and rename.col() will remove them if 
rm_col argument = TRUE)

3. decide whether a new controlled vocabulary should be added. If so, add 
to controlled vocabulary using the correct formatting, which is described in the README
(I usually get Julia's opinion on this, especially before adding new vocabularies)

```{r}
colnames_curation <- 
  read.csv(here(auxiliary_files, "colnames_summary_aux.csv")) %>%
  mutate(controlled_vocab = 
           ifelse(controlled_vocab == "", colname, controlled_vocab)) 

data_raw_rename1 <- rename.col(data_raw3, 
                              rename_df = colnames_curation, 
                              rename_col = controlled_vocab, 
                              old_col = colname, 
                              rm_col = TRUE)

save(data_raw_rename1,
     file = here(knitroutputfolder, "data_raw_rename1.RData"))
```
## Bind all data.frames
```{r}
data_all1 <- bind_rows(data_raw_rename1)
save(data_all1,
     file = here(knitroutputfolder,"data_all1.RData"))
```
## Curate column contents:

### location

## Validate trial data
See README for more information on the codebooks (Using codebooks_all_db.csv)
used to validate the data
```{r}
df_validate <- data_all1
```
### Check column names
```{r, results.folding = NULL, results = 'asis'}
colname_valid <- validate_colnames(df_validate, "trial_data") %>%
    select(comment, colname_data, colname_codebook, required, col_num) 

kable(colname_valid)

# To interact with the column names that still need to be fixed
colname_valid_check <- colname_valid %>% 
  filter(comment == "not present in codebook: trial_data")

col_info <- find_col_info(data_all1, 
                       cols_check = colname_valid_check$colname_data, 
                       by_col = sourcefile)


```

### Check column contents
```{r, results.folding = NULL, results = 'asis'}
colcontent_valid <- confront_data(df_validate, df_type = "trial_data")

colcontent_summary <- colcontent_valid[["summary"]]
kable(colcontent_summary)

# To check any fails interactively:
var <- c("rep")
colcontent_violate <- 
  violating(df_validate, colcontent_valid[[2]][var]) %>% 
  relocate(matches(var))

```

## Validate metadata
```{r}
metadata_validate <- data_all1
```
### Check column names
```{r, results.folding = NULL, results = 'asis'}
metadata_colname_valid <- validate_colnames(metadata_validate, "trials_metadata") %>%
    select(comment, colname_data, colname_codebook, required, col_num) 

kable(metadata_colname_valid)

```

### Check column contents
```{r, results.folding = NULL, results = 'asis'}
metadata_colcontent_valid <- confront_data(metadata_validate, df_type = "trials_metadata")

metadata_colcontent_summary <- metadata_colcontent_valid[["summary"]]
kable(metadata_colcontent_summary)

```

