---
title: "Introduction to facthelpeR"
author: "Jacqueline Tay"
date: "`r Sys.Date()`"
output: 
rmarkdown::html_vignette:
  fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Introduction to facthelpeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Introduction
**facthelpeR** is a set of functions that are used to read in and do a first
pass at cleaning up data.  

Load **facthelpeR**
```{r setup, message = FALSE}
library(facthelpeR)
library(here)
```
## List the file structure of the data
auxiliary_files is a folder that contains helper files that are used in the
curation process.  Generally, the *_aux.csvs are manually edited to match
controlled vocabularies.  They live in a separate folder than the output
so that they don't get written over accidentally.

```{r}
main_folder <- "inst/extdata/OSU_2019_example"
data_folder <- here::here(main_folder, "data-raw")
auxiliary_files <- here::here(main_folder, "auxiliary_files") 
knitroutput_folder <- here::here(main_folder, "output")
```
## Manual edits
If there are any mistakes in numeric values or one-off typos, sometimes it is
faster to fix it in the raw files.  Note any manual changes here in the .Rmd
for reproducibility.

## Read in raw data
### Generate a list of all of the files and sheets
If the header is not in the first row (or spans multiple rows), 
`write.csv(sheet_info, ...)` and manually fill in the 
number of rows to skip in each sheet.  This data.frame will be read into 
`read_multsheets()`
```{r}
files_csv <- list.files(data_folder, recursive = TRUE, pattern = ".csv")

sheet_info_csv <- data.frame(filename = files_csv) %>% 
  filter(!str_detect(filename, "curation_files")) %>%
  mutate(list_names = basename(filename)) %>%
  mutate(list_names = str_remove(list_names, ".csv")) 

knitr::kable(sheet_info_csv)
```
#### How to handle sheets in .xls files
This section is just to demonstrate how to handle excel files with multiple sheets.
The data won't be processed in the rest of this vignette.
```{r}
files_xls <- list.files(data_folder, recursive = TRUE, pattern = ".xls")

sheet_info_xls <- list_sheetnames(data_folder, files_xls) 

knitr::kable(sheet_info_xls)
```
### Use read.multsheets()
```{r}
data_raw1 <- read.multsheets(data_folder, sheet_info_csv, na = c("NA"), col_names = TRUE)
data_raw2 <- imap(data_raw1, function(x, y){ x %>% mutate(sourcefile = y)})


knitr::kable(head(data_raw2[[1]]))
knitr::kable(head(data_raw2[[2]]))
#save(data_raw2,
 #    file = here::here(knitroutput_folder, "data_raw2.RData"))
```
### Remove duplicate columns
Use `facthelpeR::rm_dup_col()` because it can detect and remove duplicates when:
- column names are not exact, but column contents are
- if the only difference between two column contents are that one has some NAs,
then that column is removed and the other is retained
```{r}
data_raw3 <- imap(data_raw2, function(x,y){
    print(y)
    rm_dup_col(x, FALSE) %>% 
    select(where(~!all(is.na(.))))
})

```

## Curate column names:
### Get column names
```{r}
colnames <- summarize_variables(data_raw3, reg_ex = NULL)
colnames_summary <- colnames %>%
  group_by(colname) %>%
  summarize(across(everything(), ~ paste(na.omit(unique(.x)), collapse = "; ")))

write.csv(colnames_summary, here::here(knitroutput_folder, "colnames_summary.csv"), row.names = FALSE)

knitr::kable(colnames_summary)
```
### Rename columns using auxiliary files
First, understand what each column means (description and units) by:

1. checking through annual reports

2. checking with Julia

3. lastly, asking the PI

Then, fill out the auxiliary file with the appropriate controlled vocabulary.

If there isn't a corresponding controlled vocabulary, then need to:

1. decide if the column can be captured by current controlled vocabulary. If so,
standardize to the controlled vocabulary

2. decide whether we need to keep the variable at all (can fill in auxiliary file
with "rm_col" in the controlled_vocab column and rename.col() will remove them if 
rm_col argument = TRUE)

3. decide whether a new controlled vocabulary should be added. If so, add 
to controlled vocabulary using the correct formatting, which is described in the README
(I usually get Julia's opinion on this, especially before adding new vocabularies)

```{r}
colnames_curation <- 
  read.csv(here::here(auxiliary_files, "colnames_summary_aux.csv")) %>%
  # If a column name still needs to be checked/confirmed, leave the controlled_vocab 
  # column blank in the "colnames_summary_aux.csv" file.  
  # Before the renaming step, just set the unknowns to the 
  # original name.
  mutate(controlled_vocab = 
           ifelse(controlled_vocab == "", colname, controlled_vocab)) 

knitr::kable(colnames_curation)

data_raw_rename1 <- rename_col(data_raw3, 
                              rename_df = colnames_curation, 
                              rename_col = controlled_vocab, 
                              old_col = colname, 
                              rm_col = TRUE)

knitr::kable(head(data_raw_rename1[[1]]))
knitr::kable(head(data_raw_rename1[[2]]))

save(data_raw_rename1,
     file = here::here(knitroutput_folder, "data_raw_rename1.RData"))
```
### Some functions that may help explore the columns within the list of data
There are often instances where column names are not used consistently between 
files or there are similar column names within one dataset (test vs. testwt).
The following functions may help to summarize and explore the data.

- See `summarize_variables()` used above.
- `summarize_ncol()` and `summarize_colnames` summarizes the information 
collected from `summarize_variables()` 

List the number of columns that match a given regex per dataset.
```{r}
cols_summary <- summarize_ncol(data_raw_rename1, c("entry", "length"))
knitr::kable(cols_summary)
```

List the proportion of files that contain a given 
```{r}
cols_prop <- summarize_colnames(data_raw_rename1, c("entry", "length"))
knitr::kable(cols_prop)
```

Pull all the columns that match a given regex.  This is useful for checking to 
see whether column names have been used consistently within and across datasets.
```{r}
cols <- select_colsfromlist(data_raw_rename1,  c("entry", "length"))
knitr::kable(cols)
```

## Bind all data.frames
```{r}
data_all1 <- bind_rows(data_raw_rename1)
save(data_all1,
     file = here::here(knitroutput_folder,"data_all1.RData"))
```

## Curate column contents

See the [**factcuratoR**](https://idahoagstats.github.io/factcuratoR/) package for validating columns and column contents

